Revisiting the hydrodynamic modulation of short waves by longer waves

### 2024-09-19 paper/paper.tex
Specify numerical error of centered finite difference
Numerical results

More recently, \citep{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are realistic by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario of
a uniform short-wave field that suddenly appears in a uniform long-wave train,
and the numerical simulations stabilize if the short-wave field is initialized
from existing steady solutions.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.

@@

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
the gravitational acceleration at the surface of a Stokes wave is, to the
4$^{th}$ order:

@@

To quantify the contribution of the nonlinear terms in Eqs. \ref{eq:wavenumber}
and \ref{eq:wave_action2}, and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
second-order central finite difference in space and integrated in time using the
forth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, which neglects the Stokes drift of
short-wave groups.
Their gravity modulation also does not consider the nonlinear or the curvature
effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the MUSCL4 scheme by \citet{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
We show in the next section that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
\caption{}
\label{fig:numerical_vs_analytical_k}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
\caption{}
\label{fig:numerical_vs_analytical_a}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
\caption{}
\label{fig:numerical_vs_analytical_ak}
\end{figure}

>>>>

More recently, \citet{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are realistic by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario of
a uniform short-wave field that suddenly appears in a uniform long-wave train,
and the numerical simulations stabilize if the short-wave field is initialized
from existing steady solutions.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.

@@

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
the gravitational acceleration at the surface of a Stokes wave is, to the
fourth order:

@@

To quantify the contribution of the nonlinear terms in Eqs. \ref{eq:wavenumber}
and \ref{eq:wave_action2}, and to evaluate the steadiness assumption of the
analytical solutions, we now proceed to numerically integrate the full set of
crest and action conservation equations.
Equations (\ref{eq:wavenumber})-(\ref{eq:wave_action2}) are discretized using
second-order central finite difference in space and integrated in time using the
forth-order Runge-Kutta method \citep{butcher1996runge}.
The space is divided into 100 grid points.
This is effectively the same equation set and numerical configuration as that of
\citet{peureux2021unsteady}, except that their long-wave orbital velocities are
evaluated at $z=0$ instead of $z=\eta_L$, which neglects the Stokes drift of
short-wave groups.
Their gravity modulation also does not consider the nonlinear or the curvature
effects, however, this difference does not qualitatively affect the results.
Another difference is the choice of the numerical scheme for spatial differences,
which in their case is the MUSCL4 scheme by \citet{kurganov2000new}.
Although the second-order central finite difference is not appropriate for many
numerical problems, in our case it is sufficient because it is conservative and
the fields that we compute derivatives of are smooth and continuous.
The maximum relative error of the centered finite difference in this model
is $\approx 0.066\%$.
We show in the next section that our numerical solutions for infinite long-wave
trains are qualitatively equivalent to those of \citet{peureux2021unsteady}.

@@

The first set of simulations is performed for the long-wave steepness of
$\varepsilon_L = 0.1$, $0.2$, and $0.4$ (Fig. \ref{fig:numerical_solutions}).
The long-waves are initialized from rest ($a_L = 0$) and are ramped up
to their target steepness after 5 long-wave periods, thus mimicking the
passage of a long-wave group
(we explain in more detail the importance of ramping up the long-wave forcing
in the next subsection).
At $\varepsilon = 0.1$, the numerical solutions for the wavenumber, amplitude,
and steepness modulation are all remarkably similar to the analytical solutions
(Fig. \ref{fig:numerical_solutions}a,d,g).
This suggests that at low long-wave steepness that steady analytical solution
is a reasonable proxy for the fully-nonlinear crest-action conservation
equation solutions.
The numerical solutions diverge more notably from the analytical ones at
moderate steepness of 0.2 (Fig. \ref{fig:numerical_solutions}b,e,h), and are
considerably different at very high steepness of 0.4
(Fig. \ref{fig:numerical_solutions}c,f,i).
These differences suggest that at high steepness the inhomogeneity of
gravitational acceleration and group speed become important.

@@

The maximum modulations of the short-wave wavenumber, amplitude, and steepness
from the analytical and numerical solutions, are shown on Figs.
\ref{fig:numerical_vs_analytical_k},
\ref{fig:numerical_vs_analytical_a}, and
\ref{fig:numerical_vs_analytical_ak}, respectively, as function of the long-wave
steepness $\varepsilon_L$.
Overall, the numerical solutions, based on the linear and third-order Stokes long wave
alike, begin to diverge from the analytical solution at $\varepsilon_L \approx 0.2$.
The numerical solution using the Stokes long wave is quantitatively very similar
to the fully-nonlinear solutions of \citet{longuet1987propagation} in amplitude
modulation, however with somewhat lower wavenumber and steepness modulations.
For example, at $\varepsilon_L = 0.4$, the steepness modulation in our solutions
is close to 6, whereas \citet{longuet1987propagation} finds it close to 8.
The underestimate in the wavenumber modulation at high long-wave steepness
relative to the results of \citet{longuet1987propagation}
is likely due to his long wave velocities being evaluated to a higher order
than they are here.
Nevertheless, the overall steepness modulation between the two numerical
approaches are remarkably similar, and either solution would cause the
short-waves of even low steepness to break \citep{banner1993wave}.

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_k.png}
\caption{
  Maximum wavenumber modulation as function of long-wave steepness $\varepsilon_L$.
  Black line and circles are for the analytical and numerical solutions from
  \citet{longuet1960changes} and \citet{longuet1987propagation}, respectively.
  Blue line is based on the analytical solutions from this paper.
  Green and orange lines are for the numerical solutions from this paper
  using linear and third-order Stokes long waves, respectively.
}
\label{fig:numerical_vs_analytical_k}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_a.png}
\caption{
  As Fig. \ref{fig:numerical_vs_analytical_k} but for the amplitude modulation.
}
\label{fig:numerical_vs_analytical_a}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/numerical_vs_analytical_solutions_ak.png}
\caption{
  As Fig. \ref{fig:numerical_vs_analytical_k} but for the steepness modulation.
}
\label{fig:numerical_vs_analytical_ak}
\end{figure}
